Embedding Dimension of Polyhedral Losses

### 2020-01-31 colt-20/embed-dim.tex
Uhhhh
Moving 4.4 to App
Fixing some find/replace mishaps

\begin{theorem} \label{thm:general-char}
  Let $\ell: \R \to \reals_{\geq 0}^{\Y}$ be a discrete loss with, for each $r \in \R$, $\gamma_r = \{p : r \in \argmin_{r'} \inprod{p}{\ell(r)}\}$.
  Then $\ell$ is $d$-embeddable if and only if there exists a collection of polytopes $\T^r = \{T^r_y : y \in \Y\}$ for each $r \in \R$ such that both of the following hold:
  \begin{enumerate}
    \item (Optimality) For each $r$, we have $D(\T^r) = \gamma_r$.
    \item (Monotonicity) There exists an injective embedding function $\varphi : \R \to \reals^d$ and loss functions $\{L_y : \reals^d \to \reals^\Y_{\geq 0}\}_{y \in \Y}$ such that for all $r \in \R$ and $y \in \Y$, we have $T^r_y = \partial L(\varphi(r))_y$ and for all $r \in \R$, we have $L(\varphi(r))_y = \ell(r)_y$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  $(\implies)$ 
%  If $\ell$ is $d$-embeddable, say that $L:\reals^d \to \reals^\Y_{\geq 0}$ embeds $\ell$.
%  Take $T_y$ to be $\partial L(\varphi(r))_y$ for a fixed $r \in \R$ and $y \in \Y$.
%  Let $\Gamma$ be the property elicited by the surrogate $L$.
%  To see the optimality condition is satisfied, consider that for each $r \in \R$, we have $\gamma_r = \Gamma_{\varphi(r)}$ by Bayes Risks matching~\cite[Proposition 1]{finocchiaro2019embedding} and level sets of the property mapping to subgradients of the Bayes Risk by~\cite{frongillo2014general}.
%  In turn, we have $\Gamma_{\varphi(r)}= \{p \in \simplex : \vec 0 \in \oplus_p \partial L(\varphi(r))_y \} = \{p : \vec 0 \in \oplus_p T_y \} = D(\T)$.
%  Clearly this set of polytopes $\T$ satisfies the monotonicity condition by the definition of $L$ embedding $\ell$.  
$L$ embeds $\ell$ implies $\gamma_{r} = \Gamma_\varphi(r) = D(\T^r)$ by Lemma\ref{lem:minkowski-0-opt} for all $r \in \R$, thus we have optimality.
Monotonicity is simply a corollary of the embedding definition.

@@

We now focus entirely on the optimality condition of Theorem \ref{thm:general-char}, for two purposes.
First, we aim to greatly narrow the search space for constructing low-dimensional surrogate loss functions for a given discrete loss.
The tools we construct in this section aid in this task by constraining or constructing feasible subgradient sets $\T$ given a level set $\gamma_r$.
Second, we wish to prove impossibilities, i.e., lower bounds on the embedding dimension of a given discrete loss (an apparently hard problem).
For such lower bounds, it suffices to drop monotonicity from Theorem \ref{thm:general-char}, leaving us with an independent optimality condition for each $r\in\R$, and show that for any one $r\in\R$, we could not have $d$-dimensional polytopes $\T$ satisfying $D(\T) = \gamma_r$.
% (For upper bounds, one must additionally show monotonicity.)

At first glance, the optimality condition seems difficult to operationalize, as it involves the existence of polytopes, and even if said polytopes are given, it is unclear how to test whether $D(\T)=\gamma_r$.
To begin, consider the latter problem, of understanding the set $D(\T)$ in terms of descriptions of $\T$, and in particular, of writing conditions on $\T$ such that $D(\T)$ is equal to a given polytope $C\subseteq\simplex$.
We now that, by writing $C$ in its halfspace and vertex representations, respectively, we can give two such conditions.

@@

\section{$d$-dimensional omitted proofs}\label{app:omitted-d-dim}
\begin{proof}[Proof of Lemma~\ref{lemma:minkowski-support}]
	For each $p$, $\oplus_p \T$ is a polytope.
	\bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
	For each of the finitely many supports ($2^n - 1$), we know $\oplus_p \T$ is a polytope, and every polytope can be defined by a finite, complete set of vectors for that polytope.
	As a two polytopes with the same support are combinatorially equivalent\jessiet{is this what we want?  I know the idea but not sure if this is the proper term}, they can be defined by the same facet enumeration, and any set of normals that is complete for $\oplus_p \T$ is also complete for a $\oplus_{p'}\T$ if $\supp(p) = \supp(p')$.
	We can simply concatenate these finite set of normals for the finite polytope supports, with some normals possibly becoming redundant.
	This yields finitely many normals defining the weighted Minkowski sum $\oplus_p \T$ for all $p \in \simplex$.
\end{proof}

>>>>

\begin{theorem} \label{thm:general-char}
  Let $\ell: \R \to \reals_{\geq 0}^{\Y}$ be a discrete loss with, for each $r \in \R$, $\gamma_r = \{p : r \in \argmin_{r'} \inprod{p}{\ell(r)}\}$.
  Then $\ell$ is $d$-embeddable if and only if there exists a collection of polytopes $\T^r = \{T^r_y : y \in \Y\}$ for each $r \in \R$ such that both of the following hold:
  \begin{enumerate}
    \item (Optimality) For each $r$, we have $D(\T^r) = \gamma_r$.
    \item (Monotonicity) There exists an injective embedding function $\varphi : \R \to \reals^d$ and loss functions $\{L_y : \reals^d \to \reals^\Y_{\geq 0}\}_{y \in \Y}$ such that for all $r \in \R$ and $y \in \Y$, we have $T^r_y = \partial L(\varphi(r))_y$ and for all $r \in \R$, we have $L(\varphi(r))_y = \ell(r)_y$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  $(\implies)$ 
%  If $\ell$ is $d$-embeddable, say that $L:\reals^d \to \reals^\Y_{\geq 0}$ embeds $\ell$.
%  Take $T_y$ to be $\partial L(\varphi(r))_y$ for a fixed $r \in \R$ and $y \in \Y$.
%  Let $\Gamma$ be the property elicited by the surrogate $L$.
%  To see the optimality condition is satisfied, consider that for each $r \in \R$, we have $\gamma_r = \Gamma_{\varphi(r)}$ by Bayes Risks matching~\cite[Proposition 1]{finocchiaro2019embedding} and level sets of the property mapping to subgradients of the Bayes Risk by~\cite{frongillo2014general}.
%  In turn, we have $\Gamma_{\varphi(r)}= \{p \in \simplex : \vec 0 \in \oplus_p \partial L(\varphi(r))_y \} = \{p : \vec 0 \in \oplus_p T_y \} = D(\T)$.
%  Clearly this set of polytopes $\T$ satisfies the monotonicity condition by the definition of $L$ embedding $\ell$.  
$L$ embeds $\ell$ implies $\gamma_{r} = \Gamma_\varphi(r) = D(\T^r)$ by Lemma~\ref{lem:minkowski-0-opt} for all $r \in \R$, thus we have optimality.
Monotonicity is simply a corollary of the embedding definition.

@@

We now focus entirely on the optimality condition of Theorem \ref{thm:general-char}, for two purposes.
First, we aim to greatly narrow the search space for constructing low-dimensional surrogate loss functions for a given discrete loss.
The tools we construct in this section aid in this task by constraining or constructing feasible subgradient sets $\T$ given a level set $\gamma_r$.
Second, we wish to prove impossibilities, i.e., lower bounds on the embedding dimension of a given discrete loss (an apparently hard problem).
For such lower bounds, it suffices to drop monotonicity from Theorem \ref{thm:general-char}, leaving us with an independent optimality condition for each $r\in\R$, and show that for any one $r\in\R$, we could not have $d$-dimensional polytopes $\T$ satisfying $D(\T^r) = \gamma_r$.
% (For upper bounds, one must additionally show monotonicity.)

At first glance, the optimality condition seems difficult to operationalize, as it involves the existence of polytopes, and even if said polytopes are given, it is unclear how to test whether $D(\T^r)=\gamma_r$.
To begin, consider the latter problem, of understanding the set $D(\T)$ in terms of descriptions of $\T$, and in particular, of writing conditions on $\T$ such that $D(\T)$ is equal to a given polytope $C\subseteq\simplex$.
We now that, by writing $C$ in its halfspace and vertex representations, respectively, we can give two such conditions.

@@

\section{$d$-dimensional omitted proofs}\label{app:omitted-d-dim}

\begin{proof}[Proof of Lemma~\ref{lemma:minkowski-support}]
	For each $p$, $\oplus_p \T$ is a polytope.
	\bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
	For each of the finitely many supports ($2^n - 1$), we know $\oplus_p \T$ is a polytope, and every polytope can be defined by a finite, complete set of vectors for that polytope.
	As a two polytopes with the same support are combinatorially equivalent\jessiet{is this what we want?  I know the idea but not sure if this is the proper term}, they can be defined by the same facet enumeration, and any set of normals that is complete for $\oplus_p \T$ is also complete for a $\oplus_{p'}\T$ if $\supp(p) = \supp(p')$.
	We can simply concatenate these finite set of normals for the finite polytope supports, with some normals possibly becoming redundant.
	This yields finitely many normals defining the weighted Minkowski sum $\oplus_p \T$ for all $p \in \simplex$.
\end{proof}
